---
title: "Predicting MA132 (Calculus II) Enrollment"
author: Kyle Monette
output: html_notebook
---

Can we predict spring enrollment for MA132 using the enrollment data for MA131 (Calculus I) from the prior semester?

```{r}
library(tidyverse)
library(lubridate)
library(glue)
library(scales)
library(ggrepel)
library(DT)
```

Read in the data.

```{r}
df1 <- read_csv("ma132.csv", col_types = cols(`Acad Org` = col_skip(), `Current Status` = col_skip(), `Course ID` = col_skip(), Term...7 = col_skip(), Instructor = col_skip(),'Title'=col_skip(),'Section'=col_skip()), skip = 1) %>%
  rename(Term=Term...8)
```

Filter to MA131/132 and compute enrollment.

```{r}
class_data <- df1 %>%
  filter(Catalog %in% c(131,132),Component=="LEC") %>%
  select(Catalog,Term,`Tot Enrl`) %>%
  group_by(Catalog,Term) %>%
  summarise(Enroll = sum(`Tot Enrl`)) %>%
  ungroup()
```

Split apart the term into fall/spring and the year.

```{r}
df2 <- class_data %>%
filter(Term != "Summer") %>%
separate(Term,c("Term","Year")) %>%
select(Catalog,Term,Year,Enroll) %>%
mutate(
  calc1 = case_when(Catalog == 131 ~ c(Enroll)),
  calc2 = case_when(Catalog == 132 ~ c(Enroll))
) 
```

Convert the year and term into an academic year format.

```{r}
df3 <- df2 %>%
select(Term,Year,calc1,calc2) %>%
mutate(Year = as.numeric(Year)) %>%
mutate(Year = Year - 2000) %>%
mutate(ay = case_when(Term == "Fall" ~str_c(Year, "-", Year + 1),
Term == "Spring" ~str_c(Year - 1, "-", Year)))
```

Collapse the data vertically, grouped by term and year.

```{r}
df4 <- df3 %>%
  select(Term,ay,calc1,calc2) %>%
  group_by(Term,ay) %>%
  summarise_all(~first(na.omit(.))) %>%
  ungroup()
```

Make columns for enrollment based on each fall/spring and calc1/calc2 combination.

```{r}
df5 <- df4 %>%
  group_by(ay) %>%
  mutate(
    calc1f = case_when(Term == "Fall" ~ c(calc1)),
    calc1s = case_when(Term == "Spring" ~ c(calc1)),
    calc2f = case_when(Term == "Fall" ~ c(calc2)),
    calc2s = case_when(Term == "Spring" ~ c(calc2))
  ) %>%
  ungroup()
```

Collapse the data vertically again to get the format of academic year with each enrollment.

```{r}
df6 <- df5 %>%
  select(ay,calc1f,calc1s,calc2f,calc2s) %>%
  group_by(ay) %>%
  summarise_all(~first(na.omit(.))) %>%
  ungroup()
```

Plot the results.

```{r}
ggplot(df6, mapping=aes(x = calc1f, y = calc2s)) + geom_point() + geom_smooth(method='lm') + xlab("Calculus I Fall") + ylab("Calculus II Spring")
```
